Pool Model Paper

### 2023-08-08 pool_model.tex
small changes
add citations

Consider for simplicity the system where only one bacteria species is present $N=1$.
According to the growth curves (\cmt{some references on bacteria growth curves??}) quite often at the beginning no increase in bacterial count is present and so-called lag-phase is observed.
This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
After this some bacteria enter the exponential growth phase, while some of it may not be able to adjust and experience the cell death.
Some bacteria even may not be able to adjust and die other will enter the exponential growth phase after a lag phase. 
After exponential growth the system reaches the saturation where the balance between the number of dying bacteria is in balance with the growing one leads to the stationary state.
In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.
Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{SchematicRep}}): 
\begin{itemize}
\item $L(t)$: the fraction of the population in the lag phase
\item $G(t)$: the fraction of the population growing and dividing
\item $D(t)$: the fraction of the population undergoing cell death
\end{itemize}
Here the concentrations of the pools are components of our variable vector $\mbx (t) =\{L(t), G(t), D(t)\}$.

@@

Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource. \cmt{ref review}
In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient. \cmt{ref Monod}
\cmt{Ref about Quorum sensing. Signalling mechanism that controll gene expression (bacteria growth within species) in response to environment. }

@@

As it was mentioned above, bacteria usually coexist in large multispecies interacting communities whose growth is vastly determined by the developed survival mechanisms.
These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones.
Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species.

@@

One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio. \cmt{Ref Tilman 1977, Ref Smith 2002}
To descibe such a situation we consider for simplicity only one common limiting resource pool. 
The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):

@@

which has the structure of a generalised Lotka-Volterra system. In order to better see this, we set $L_A=L_B=0$: 
\begin{eqnarray}
\label{Comp_Model}
\dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right ) - \frac{\alpha_A}{N_t}G_AG_B\\
\dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right ) -\frac{\alpha_B}{N_t}G_AG_B. 
\end{eqnarray}

@@

@@

@@

>>>>

Consider for simplicity the system where only one bacteria species is present $N=1$.
According to the growth curves (\cmt{some references on bacteria growth curves??}) \cite{buchananLifePhasesBacterial1918, rolfeLagPhaseDistinct2012} quite often at the beginning no increase in bacterial count is present and so-called lag-phase is observed.
This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
\cmt{\cite{rolfeLagPhaseDistinct2012}: About Lag-phase machanisms.}
After this some bacteria enter the exponential growth phase, while some of it may not be able to adjust and experience the cell death.
Some bacteria even may not be able to adjust and die other will enter the exponential growth phase after a lag phase. 
After exponential growth the system reaches the saturation where the balance between the number of dying bacteria is in balance with the growing one leads to the stationary state.
In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.
Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{SchematicRep}}): 
\begin{itemize}
\item $L(t)$: the fraction of the population in the lag phase
\item $G(t)$: the fraction of the population growing and dividing
\item $D(t)$: the fraction of the population undergoing cell death
\end{itemize}
Here the concentrations of the pools are components of our variable vector $\mbx (t) =\{L(t), G(t), D(t)\}$.

@@

Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource. \cmt{ref review}
In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient. \cmt{ref Monod}
\cite{monodGrowthBacterialCultures1949}
\cmt{Ref about Quorum sensing. Signalling mechanism that controll gene expression (bacteria growth within species) in response to environment. \cite{ngBacterialQuorumsensingNetwork2009}}

@@

As it was mentioned above, bacteria usually coexist in large multispecies interacting communities whose growth is vastly determined by the developed survival mechanisms.
These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones. \cite{hibbingBacterialCompetitionSurviving2010, stubbendieckBacterialCommunitiesInteractions2016} \cmt{Both good reviews on bacterial interaction}
Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species.

\cmt{\cite{h.tusonBacteriaSurfaceInteractions2013}: Attachment of bacteria to surfaces. (One of the mechanisms to stick to the areas with nutritients)}

@@

One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio. 
\cmt{Ref Tilman 1977 \cite{tilmanResourceCompetitionPlankton1977}, Ref Smith 2002 \cite{smithEffectsResourceSupplies2002}}
To descibe such a situation we consider for simplicity only one common limiting resource pool. 
The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):

@@

which has the structure of a generalised Lotka-Volterra system. \cmt{Add reference about Lotka-Volterra. Compare with different states LV.}
In order to better see this, we set $L_A=L_B=0$: 
\begin{eqnarray}
\label{Comp_Model}
\dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right ) - \frac{\alpha_A}{N_t}G_AG_B\\
\dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right ) -\frac{\alpha_B}{N_t}G_AG_B. 
\end{eqnarray}

@@

\cmt{\cite{lenskiCoexistenceTwoCompetitors1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}

@@

\cmt{\cite{rochaBaseCompositionBias2002}: Symbiosis host/bacteria. bacterial genomes suffering reductive evolution due to this. Maybe it's connected with 'social cheaters'?}

@@

\bibliography{references}
